---
title: "The Report About Incarcentration in the United State"
author: "Bella Ge"
date: "11/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r include = FALSE}
library("knitr")
```

```{r include = FALSE}
source("analysis.R")
```

## Introduction

The **problem domain** in this report is comparing the disproportionate incarceration between black people and white people in jail in the United State. As we know, there is a difference in the proportion of different people of color in jail. Therefore, I am curious about which county in the United State has the most severe issue of disproportion in black and white people in jail. 

The **measure** of incarceration I am going to focus on is the jail incarceration.

The **variables** I am going to use in this analysis are _**years, county_name, white_jail_pop, black_jail_pop, total_jail_pop.**_ I decides to use white_jail_pop, black_jail_pop, and total_jail_pop, because I want to clearly show the discrepancy in the population count of white people and black people in jail. Also, I select years because I need to use it to track the **changes** in the population count of white people and black people in jail. I choose county_name as the fifth variable since I want to find which county has **the most disproportionate incarceration** between black people and white people.

## Summary Information
 Below, there are eight values that I calculated using DPLYR skills. With the variables that I choose, I firstly can get the year when the highest jail population occurred in the United State, which is **1993**. In 1993, the number of jail incarceration reached its highest point. In this year, the population of black people and white people also reached the highest point. There were about **13143** black people and about **7036** white people in jail in 1993. Plus, the average number of black people in jail is about **74** and the average number of white people in jail is about **65**. From these simple data, we can clearly find the pattern that the numbers of black people is twice as the numbers of white people in jail. Besides these discoveries, I also found that **Los Angeles County** has both the highest number of black people as well as the highest number of white people in jail. in addition, I discovered the ratio between black people and white people in jail in Las Angegles County in 2018, which is **1.948042**. From this data, it also proved that the population of black people who are in jail is still double than the population of white people who are in jail in recent years. Based on this evidence, I can assume that the proportionate incarceration is still happening in current jail system and the disproportion is affected by the difference in race since black people get double chance to get into the jail than white people.


 * The year which has the *highest jail population count* in the United State
```{r echo=FALSE}
high_year 
```
* The average number of *Black people* in the jail incarceration across all the counties in 1993
```{r echo=FALSE}
mean_black_1993
```
* The average number of *White people* in the jail incarceration across all the counties in 1993
```{r echo=FALSE}
mean_white_1993
```
* The highest numbers in *Black people who are in jail* 
```{r echo=FALSE}
max_black_jail
```
* The highest numbers in *White people who are in jail*
```{r echo=FALSE}
max_white_jail
```
* The county has the highest number of *Black people who are in jail*
```{r echo=FALSE}
county_high_black
```
* The county has the highest number of *White people who are in jail*
```{r echo=FALSE}
county_high_white
```
* The ratio of *Black people who are in jail* to *White people who are in jail* in Los Angeles County in 2018
```{r echo=FALSE}
ratio_black_white
```

## Trends Over Time Chart
```{r echo=FALSE}
trend_chart
```

I want to include this chart because I want to discover the trends of the population 
count of black people who are in jail in the top two counties which has the most jail 
population count. I want to know whether there is an increased tendency in the population 
count of black people in jail because of the impact of racism. 
This chart displays that the number of black people who are in jail has a peak from 1990
to 1995. However, the overall trend of the population count is dramatically decreasing 
after the peak from 1990 to 2020 in Los Angeles County and New York County, which implies
that the inequality in the jail is eased now. 

## Variable Comparison Chart
```{r echo=FALSE}
comapre_chart
```

I made this chart because I want to compare the population count of black and 
white people who are in jail in Los Angeles County that has the biggest jail 
population. I want to show the specific disproportion of number between black 
people and white people. 
This chart displays the comparison of the population count between black people 
and white people. From the chart, we can see the the population of black people 
is always double than the population of white people in jail. For example, when 
the population count of black people is 8000, the population count of white people
is about 4500. Based on this evidence, it is clearly to understand why there is 
a disproportion in the number of black people who are in jail and white people who
are in jail. And, I can suggest that black people bear more social pressure and 
experience more inequality in society since they are more likely to be caught in 
prison.


## Map
```{r echo=FALSE}
Amercia_map
Amercia_map_2
```

The reason why I include this chart is that I want to compare the distribution of
white and black people who are in jail in the United State in 2013. I also want to 
display which part of the United State contains the highest number of prisoners.
From these two maps, we can notice that the highest number of black and white people
who are in jail is located in the Southwest part of the United State, especially 
in California. And, in the middle part of the east of the United State, the number of
black and white people who are in jail are both the fewest. Based on these findings, the
patterns of distribution of black and white people in jail are pretty similar in 2013
even though the number of black people in jail are twice the number of white people
in jail.